---
title: "Homework 1"
author: "Sidney Gerst"
date: "January 17, 2020"
output: html_document
---
<style type="text/css">

h1.title {
  text-align: center;
  font-family: "Times New Roman", Times, serif;
}
h4.author {
  font-family: "Times New Roman", Times, serif;
  text-align: center;
  font-size: 26px;
}
h4.date {
  font-family: "Times New Roman", Times, serif;
  text-align: center;
}

body {
font-family: "Times New Roman", Times, serif;
font-size: 20px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


  1. Three reasons why plants or animals might be patchily distributed:
       + Reason 1
       + Reason 2
       + Reason 3

###**Eureka Dune Grass**

  2. 
```{r message = F }

#Read in Swallenia Data
swallenia <- read_csv("Swallenia.csv")

#Perform two sample t-test
swallenia_ttest <- t.test(swallenia$count_2009, swallenia$count_2010, var.equal = TRUE)

swallenia_ttest


```

Based on an $\alpha$ = 0.05 the difference in mean abundance of swallenia does not have a significant change between 2009 and 2010 (p = `r round(swallenia_ttest$p.value, 2)`). This alpha is appropriate because it is a 5% risk of concluding a difference exists when there is no actual difference. 

  3. 
```{r message = F}

#Paired t-test
swallenia_pairedtest <- t.test(swallenia$count_2009, swallenia$count_2010, paired = TRUE)

swallenia_pairedtest

```

The true difference in means of abundance of swallenia between 2009 and 2010 does not equal to 0, so there is a difference in abundance (p = `r round(swallenia_pairedtest$p.value, 2)`). 

  4. 

The first analysis is more correct because (LOOK UP DEFINITION OF PAIRED TEST)

  5. 

Paragraph about why

###**Yellowstone Grizzly Bears**
  
  6. 
```{r message = F}

grizzly <- read_csv("grizzlydata.csv")




```

