---
title: "Homework 1"
author: "Sidney Gerst"
date: "January 17, 2020"
output: html_document
---
<style type="text/css">

h1.title {
  text-align: center;
  font-family: "Times New Roman", Times, serif;
}
h4.author {
  font-family: "Times New Roman", Times, serif;
  text-align: center;
  font-size: 26px;
}
h4.date {
  font-family: "Times New Roman", Times, serif;
  text-align: center;
}

body {
font-family: "Times New Roman", Times, serif;
font-size: 16px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


  1. Three reasons why plants or animals might be patchily distributed:
       + Plants don't have the ability to move away. They may drop seeds that fall directly by where they live forcing themselves to be clumped or patchy. 
       + Animals tend to live with their own species for protection such a herd of american bison. 
       + Suitable habitat is essential for an ecosystem. Nutrient concentration may be a factor for plants where patched soil concentrations effect where the plants grow. Animals will follow that pattern as grazers will feast on the plant patches and carnivores will stay close to their prey. 

**Eureka Dune Grass**

  2. 
```{r message = F }

#Read in Swallenia Data
swallenia <- read_csv("Swallenia.csv")

#Perform two sample t-test
swallenia_ttest <- t.test(swallenia$count_2009, swallenia$count_2010, var.equal = TRUE)

swallenia_ttest


```

Based on an $\alpha$ = 0.05 the difference in mean abundance of swallenia does not have a significant change between 2009 and 2010 (p = `r round(swallenia_ttest$p.value, 2)`). This alpha is appropriate because it is a 5% risk of concluding a difference exists when there is no actual difference. 

  3. 
```{r message = F}

#Paired t-test
swallenia_pairedtest <- t.test(swallenia$count_2009, swallenia$count_2010, paired = TRUE)

swallenia_pairedtest 

```

The true difference in means of abundance of swallenia between 2009 and 2010 does not equal to 0, so there is a difference in abundance (p = `r round(swallenia_pairedtest$p.value, 2)`). 

  4. 

The second analysis with the paired t-test is more accurate than a two sample t-test because each plot is not independent of eachother between the two years.  

  5. 



**Yellowstone Grizzly Bears**
  
  6. 
```{r message = F}

grizzly <- read_csv("grizzlydata.csv") 

#Filter years 1959 to 1968
grizzly5968 <- grizzly %>%
  filter(Year >= "1959" & Year <= "1968")


#perform linear regression
grizzly5968_model <- lm(N ~ Year, data = grizzly5968)

grizzly5968_model
# N(grizzlies) = 1543 - 0.7636(Year)

#Visualize the data
girzzly5968_plot <- ggplot(grizzly5968, aes(x = Year, y = N)) +
  geom_point() +
  geom_smooth(method = lm, se = T, size = 0.5)+
  theme_bw() +
  scale_x_continuous(expand = c(0,0), limits = c(1958, 1970)) +
  labs(x = "Year", y = "Number of Grizzlies")


girzzly5968_plot

```

