---
title: "Homework 2"
author: "Sidney Gerst"
date: "January 27, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

####1. Analyze the Model

```{r}

#create the model
r <- 0.05
K <- 2000
P <- 4
a <- 0.05
h <- 0.2
function1 <- function(N) (r*N)*(1-(N/K))-((P*a*N)/(1+(a*h*N)))

#plot it
function1_plot <- ggplot(data.frame(N = 0:2000), aes(x = N)) +
  stat_function(fun = function1, color = "red") +
  geom_hline(yintercept = 0) +
  theme_bw()+
  ylab("dN/dt")

function1_plot


```

#### 2. How many equilibria? 

There are three equilibira in this model (0,0), (375,0), and (1500,0). The points (0,0) and (1500,0) are stable and the point (375,0) is unstable. 

#### 3. 

When N = 0, the intrinsic population growth rate $\frac{dN}{dt}$ = 0, and is a equililbium. 

To evaluate N at equilibium, the derivative should be calculated: 

$\frac{df(N)}{dN} = r(1-\frac{2N}{K}) - P(\frac{a}{(1 + ahN)^2})$ 

When N = 0, the equation becomes: $\frac{df(N)}{dN} = r-P*a$

Then we solve for P: $P = \frac{r}{a}$ . Hunters = the growth rate per prey attack rate. P needs to be greather than this number

where is P stable, when it's larger than r/a or less than

need a number of hunters that's larger than what? 


####4. Non-zero equilibria

$0 = r*N*(1-\frac{N}{k})-P*(\frac{aN}{1 + ahN})$

####5. Graph model

A. 

```{r}

r <- 0.05
K <- 2000
P <- 4
a <- 0.05
h <- 0.2

function2 <- function(N) (r*N)*(1-(N/K))
function3 <- function(N) ((P*a*N)/(1+(a*h*N)))

function2_plot <- ggplot(data.frame(N = 0:2000), aes(x = N)) +
  stat_function(fun = function2, color = "red") +
  stat_function(fun = function3, color = "blue") +
  geom_hline(yintercept = 0) +
  theme_bw()+
  ylab("dN/dt")

function2_plot



```

B. 

When the intrinsic growth rate is greater than the hunting rate, the population will continue to grow. When the hunting rate is greater than the intrinsic growth rate, the population will go down to 0. When they are equal, the population will be at equilibrium. 

C. 

All three equilibrium points are at the same N values as in Problem 1. 

####6. 

```{r}
########## P = 1
r <- 0.05
K <- 2000
P <- 1
a <- 0.05
h <- 0.2

function4 <- function(N) (r*N)*(1-(N/K))
function5 <- function(N) ((P*a*N)/(1+(a*h*N)))

function3_plot <- ggplot(data.frame(N = 0:2000), aes(x = N)) +
  stat_function(fun = function4, color = "red") +
  stat_function(fun = function5, color = "blue") +
  geom_hline(yintercept = 0) +
  theme_bw()+
  ylab("dN/dt")

function3_plot




```

There are two equilibria. An unstable equilibrium at N = 0, and a stable at N ~ 1900. 

```{r}
############ P = 6

r <- 0.05
K <- 2000
P <- 6
a <- 0.05
h <- 0.2

function6 <- function(N) (r*N)*(1-(N/K))
function7 <- function(N) ((P*a*N)/(1+(a*h*N)))

function4_plot <- ggplot(data.frame(N = 0:2000), aes(x = N)) +
  stat_function(fun = function6, color = "red") +
  stat_function(fun = function7, color = "blue") +
  geom_hline(yintercept = 0) +
  theme_bw()+
  ylab("dN/dt")

function4_plot



```

There is only one equilibrium at N = 0 and it is stable. 

####7. Bistability

A. 

The domain of attraction for the zero equilibrium goes from 0 - 400. The domain of attraction for the largest equlibrium is any number greater than 400. 

B. 

The hunting rate is greater than the intrinsic growth rate, so you would be able to extirpate the invasive species.

C. 

When the population is at the carrying capacity and the harvest rate is greater, the population will come back to about 1500 individuals.  


#### 8. Derivitives 

A. 

```{r}

r <- 0.05
K <- 2000
a <- 0.05
h <- 0.2
d <- 1/(a*h)

function8 <- function(P) (1/2) * ((K - d) - sqrt((K-d)^2 + ((4*K*d)/(r)) * (r - a * P)))

function9 <- function(P) (1/2) * ((K - d) + sqrt((K-d)^2 + ((4*K*d)/(r)) * (r - a * P))) 

function_8_9_plot <-ggplot(data.frame(P = 0:7) , aes(x = P)) +
stat_function(fun = function8, color = "red") + 
  stat_function(fun = function9, color = "blue") +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 1, linetype = "dotted", color = "purple", size = 2) +
  theme_bw() +
  ylab("N")

function_8_9_plot


```

B. make a vertical line at the value of p where the value goes from unstable to stable.

The equilibium is unstable at P = 1. 

